--- 
title: "The Use of Cost Survey Data to Calculate Vessel Profits for Commercial Fishing Fleets in the Northeast U.S."
author: "Greg Ardini, Samantha Werner, Elizabeth Conley"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
---

# Introduction

Profit is an important metric in assessing the economic performance of commercial fishing fleets and in evaluating the impacts of proposed regulations. Evaluation of economic impacts is required under a number of laws, including Including the Magnuson-Stevens Fishery Conservation and Management Act (MSA), the National Environmental Policy Act (NEPA), Executive Order 12866, and the Regulatory Flexibility Act (RFA). 

Within the Northeast United States, commercial revenues are readily available to NOAA Fisheries, as all federally permitted vessels are required to report sales to licensed dealers. Cost data, on the other hand, have generally been the limiting factor in calculating vessel profits. The Northeast Fisheries Science Center conducts a voluntary survey to collect comprehensive cost information from commercial fishing vessel owners. This survey was last implemented during 2023 for costs incurred during calendar year 2022. One of the primary goals of the survey was to calculate vessel-level profits. 

Using data primarily collected from the most recent cost survey, this site presents commercial fishing vessel profits during 2022 for those vessels who participated in the survey effort. All information presented follows data confidentiality rules as required under the MSA (at least 3 vessels must be present in each summary group). These snapshots will serve as a performance metric in evaluating the economic impacts of proposed fishery management actions. 

## Cost Survey Background

Starting in 2007, the Social Sciences Branch (SSB) of the Northeast Fisheries Science Center (NEFSC) has intermittently collected commercial fishing cost information from vessel owners through a voluntary survey. The most recent survey, formally known as the Greater Atlantic Region Commercial Fishing Business Cost Survey for 2022 (hereafter 2022 cost survey), was fielded in March-August 2023 to collect costs incurred by vessel owners during calendar year 2022. The 2022 cost survey was split up into the following sections:

  1. General Information
  2. New/Replacement, Repair/Maintenance, Upgrade and Other Costs for the Selected Vessel
  3. Trip Expenses for the Selected Vessel
  4. Crew Payment System for the Selected Vessel
  5. Overhead Costs
  6. The Value of the Selected Vessel
  7. Quota Costs and Revenues for the Selected Vessel
  8. Completeness of Cost Information
  9. Survey Follow-up Questions

Each owner of a federally permitted commercial fishing vessel that was active in either calendar year 2021 or 2022 received a survey. A total of 2,495 individuals were sent the 2022 cost survey with 422 surveys returned. After accounting for incomplete returned surveys and ineligibles within the sampling frame, the final response rate was 15.0% (371/2,477). For further details on the implementation of the 2022 cost survey see Werner et al. (2025). For more information on early iterations of the cost survey, see Das (2015) and Ardini et al. (2022). An overview of survey data collected, as well as applications for cost data, can be found on the Cost Survey Landing Page. Additionally, the Cost Data Visualization Tool presents distributions for the various costs provided by survey respondents.


# Metrics

There are a number of different metrics to measure net returns to commercial fishing businesses as detailed in Kitts et al. (2022). For example a cash flow statement typically shows returns over time and can help inform the need to acquire cash reserves through loans. An income statement is somewhat similar to a cash flow statement and focuses on gross revenues and costs associated with operating a fishing business. An economic profit statement includes opportunity costs, such as physical capital costs and the vessel owner’s time. A vessel may yield a positive accounting profit but a negative economic profit after accounting for opportunity costs.

This page provides modified income statements looking at both operating profit and total profit. Operating profit can help inform the immediate viability of a fishing business- if trip expenses cannot be covered, economic theory suggests that the vessel will cease operations. Total profit measured over an extended time period informs the long-term viability of a fishing business. Measured for 2022, this metric can be used to calculate profit margin, the ratio of profit to gross revenues. However, a negative total profit for 2022 does not necessarily translate into a fishing business that is not viable as some costs, such as vessel repairs, are expected to fluctuate from year to year.

## Operating Profits
Operating profit refers to gross revenues minus operating costs (variable costs). Within the 2022 cost survey, two sections in particular cover operating costs. The trip expenses section (#3) collects information on fuel, bait, ice, and food. The crew payment section (#4) includes the total amount paid to crew and hired captains, as well as any benefits paid out. Outside of the survey, a few costs collected by at-sea observers, fresh water and crew supplies, are also included in the operating profit calculation.

EQUATION 1: Gross Revenue - (Trip Expenses + Crew Payments)

## Total Profits
Total profit refers to operating profit minus non-operating costs (fixed and quasi-fixed costs). Within the 2022 cost survey, a few different sections include non-operating costs. The vessel repairs, upgrades, and other costs section (#2) includes the amount spent on various vessel components including the engine, hull, fishing gear, electronics, and safety equipment. Other vessel costs such as mooring fees, permit fees, and vessel insurance are also included. The overhead section (#5) includes, among other costs, business vehicle expenses, association fees, non-crew labor payments, and office/storage expenses. The quota costs and revenue section (#7) is also included as a component of non-operating costs.

EQUATION 2: Operating Profit - (RMUI/other Vessel Costs  + Overhead + Quota Leasing)


# Respondent Pool

A total of 422 surveys were returned to the Social Sciences Branch at the conclusion of the 2022 cost survey. Profit calculations were first performed for all observations followed by a series of steps to eliminate those observations that were deemed inappropriate to include.

## Filtering of Observations
The first filtering step specified that all surveys must have answered at least one cost-related question. This step eliminated 55 observations, bringing the total number to 367. The second filtering step specified that the responding vessel must be commercially active (revenues >$0) during 2022. A further 46 observations were dropped, bringing the total to 321.

Following completion of the initial filtering, a few other steps were performed. The 2022 cost survey included the nine sections listed under Cost Survey Background. For each response, the number of survey sections skipped were counted, where a skipped section was any case in which all questions within the section were left blank. For this exercise, five survey sections were evaluated. Sections 3-5 were evaluated in entirety while section 2 was divided into sub-sections of vessel repair/maintenance/upgrade/improvement (RMUI) questions and vessel fees/insurance questions. If a survey skipped multiple sections, it was removed from the dataset. A total of 17 observations skipped multiple survey sections, bringing the total to 304. An additional 12 observations out of the remaining 304 skipped one survey section. These cases were further evaluated to look at the specific section skipped. Four out of the 12 only skipped the crew payment section. Since these were all small vessels with relatively low costs and operating profit, assuming there was no crew on board was deemed appropriate. The other 8 observations that skipped one survey section were dropped, resulting in a final dataset of 296 observations. Out of a possible 31 questions answered, these observations provided responses to 28 questions on average, with a range of 20-31 questions answered. 


# Profit Calculation Methods

## Commercial Revenue
All individuals that were sampled for the 2022 cost survey were owners of federally-permitted commercial fishing vessels active in 2021 or 2022. Federally-permitted vessels are required to sell their catch to federally-licensed seafood dealers and commercial revenue is derived from harvested catch on commercial trips sold to a licensed dealer. All vessels included in the profit snapshots generated commercial fishing revenue >$0 during 2022 from landings occurring in the Northeast region (ME to Cape Hatteras, NC). Of note, the reporting requirements for landings of bluefin tuna differ from other species landed in the Northeast region. Bluefin tuna revenue was queried from the Atlantic Coastal Cooperative Statistics Program (ACCSP) data warehouse, while all other species revenue was queried from the Catch Accounting and Management System (CAMS) housed at the Greater Atlantic Regional Fisheries Office (GARFO) and the NEFSC.

### Primary Gear
Vessel primary gear is defined according to the largest contributor to total commercial vessel revenue during 2022. Fishing gears, primarily from Vessel Trip Reports (VTRs) were grouped into the same categories as used for the 2022 cost survey sampling frame. For example, bottom trawl and midwater trawl were grouped into a broader trawl category. The resulting primary gear groups are as follows: gillnet, handgear, hydraulic dredge, longline, pot/trap, scallop dredge, seine, trawl, and unknown. Due to low response levels for some primary gear groups, results are unable to be presented.

### Primary Fishery
Vessel primary fishery is also defined as the largest contributor to total commercial vessel revenue during 2022. Species were grouped into Fishery Management Plans (FMPs) as managed by the New England and Mid-Atlantic Fishery Management Councils, as well as American lobster, managed by the Atlantic States Marine Fisheries Commission. All other fisheries are grouped into “Other”, which covers a variety of primary species/fisheries including barracuda, bay scallop, cancer crab, eastern oyster, horseshoe crab, cutlassfish, king mackerel, menhaden, ribbonfish, shrimp, tautog, and whelk. Due to low response levels for some primary fisheries, results are unable to be presented.

### Primary Gear/Fishery
Vessel primary gear (4.1.1) and primary fishery (4.1.2) were also combined to examine differences within more specific fleets. For most gear types, multiple fisheries are represented. However, scallop dredge gear, for example, is represented by only one fishery (sea scallop). Due to low response levels for some primary gear/fisheries, results are unable to be presented.

## For-Hire Revenue
While the 2022 cost survey was sent to vessels engaged in commercial fishing during 2021 or 2022, some of these vessels also were engaged in for-hire fishing. In order to most accurately capture vessel-level profits, it is important to include this additional source of revenue. For-hire revenue is derived from angler/customer payments on recreational trips.

All 2022 party/charter trips, and the number of anglers per trip, were first pulled from logbook data. Average angler charter/guide fees by state were then retrieved from the results of the 2022 National Angler Expenditure Survey administered by NOAA Fisheries. This survey is conducted every 3-5 years to better understand the economic contribution of recreational fishing within each region. The survey collects information on a variety of angler expenditures including bait, ice, lodging, and travel costs. For the purpose of the vessel profit calculation, only the amount paid by the angler for the party/charter ticket is used. State-level averages from the angler expenditure survey were used in light of the variability in fees by state within the Northeast region as well as the high level of response for each state. The average angler fee was multiplied by the number of anglers for each 2022 party/charter trip by state. Finally, the trip-level calculations were summed to an annual vessel-level value.

## Survey Costs
The vast majority of cost information used in the calculation of vessel profits came directly from the 2022 cost survey. Some manipulation of the data was required to make it usable at the vessel/annual level. These data manipulation steps are summarized below.

Operating Costs: Survey respondents were given the option of providing operating costs such as fuel, bait, and ice either at the annual level or as an average by trip. For those respondents who filled in the average cost by trip, the value was multiplied by the number of commercial trips taken during 2022. In general, the number of trips was used directly from the survey (Q2). In cases where this information was not filled in, the number of trips was retrieved from VTRs and/or dealer reports.

Overhead Costs: Survey respondents were given the option of providing these costs either at the vessel level or at the business level. For those respondents who filled in business level costs, the value was divided by the number of vessels owned by the respondent during 2022. The number of vessels owned was taken directly from the survey (Q13).

Quota Leasing Costs: Survey respondents were given the option to provide quota leasing costs either at the vessel-level or across multiple vessels. If this cost was provided across multiple vessels, the value was divided by the specific number of vessels the quota leasing cost was attributed to (Q17).

Other Costs: At the end of the survey, respondents were given the opportunity to list other costs that were not included in the survey. When other costs were listed, they were evaluated on a case-by-case basis (see Werner et al. 2025). Those that could not be allocated to an existing cost category were excluded from the profit calculation.

## Additional Trip Costs 
As covered in the previous section, the 2022 cost survey queried a variety of expenses faced by vessel owners. During pre-testing of the survey instrument, fishing industry members generally felt that the survey did a good job of including all significant costs. Furthermore, comments on the survey itself and a lack of other costs filled in by respondents indicated the survey covered significant costs. However there are two costs, crew supplies and fresh water, that are collected by at-sea observers that were not included in the 2022 cost survey. 

To capture these costs, data from observer programs was used to calculate average costs per trip by gear. All 2022 trips from observer data tables were pulled, along with the primary gear utilized on each trip. The average crew supply and fresh water trip costs by gear type were calculated and these averages were then merged by gear with all of the trips taken by survey responding vessels during 2022. Finally, trip-level costs were summed to the vessel-level.

## Depreciation
The final cost component included in the profit calculation is vessel depreciation. Ideally, a different depreciation value would be assigned to each physical capital component (e.g. electronics, engine, hull) using the estimated market value and age of each piece of vessel equipment. This method would require additional information beyond that collected in the 2022 cost survey. 

Information on the estimated market vessel value, including all equipment and fishing gear, was collected in the 2022 cost survey. However the number of responses to this question were lower than most other survey questions explained in part by the inclusion of a “I Prefer not to Answer” option, which was not included for other survey questions. Given the relative lack of response for vessel market value, an alternate method was used in which shadow values of vessel characteristics- age, gross tonnage, length, and type of construction were estimated. The shadow values were then multiplied by each vessel characteristic to calculate a total vessel value. From the vessel values, a depreciation cost was calculated as 4% of the value for those vessels less than 25 years old and 2% of the value for those vessels 25 years or older.


# Revenue Summary

```{r, echo=FALSE}
library(dplyr)
library(ggplot2)

gear_revenue<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/revenue_summary/gear_revenue_no_conf.csv")
gear_revenue_2 <- gear_revenue %>% mutate(Mean = prettyNum(mean_rev, big.mark = ',', scientific = FALSE))
gear_revenue_3 <- gear_revenue_2 %>% mutate(Median = prettyNum(med_rev, big.mark = ',', scientific = FALSE))
gear_revenue_4 <- gear_revenue_3 %>% mutate(St.Dev = prettyNum(std_rev, big.mark = ',', scientific = FALSE))
gear_revenue_4$Mean<-paste("$",gear_revenue_4$Mean)
gear_revenue_4$Median<-paste("$",gear_revenue_4$Median)
gear_revenue_4$St.Dev<-paste("$",gear_revenue_4$St.Dev)
gear_revenue_4$Response_Perc<-paste(gear_revenue_4$Response_Rate,"%")
gear_revenue_4 <- subset(gear_revenue_4, select=c(Gear, Observations, Sampled, Response_Perc, Mean, Median, St.Dev))
knitr::kable(gear_revenue_4)
gear_revenue_plot<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/fleets/gear_boxplots.csv")
gear_revenue_plot_2 <- ggplot(gear_revenue_plot, aes(x = Gear, y = total_revenue, fill=Gear)) + geom_boxplot(show.legend = FALSE, outlier.shape = NA)
gear_revenue_plot_2 + coord_cartesian(ylim=c(0, 2000000)) + scale_y_continuous(breaks = seq(0, 2000000, by=250000))

fmp_revenue<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/revenue_summary/fmp_revenue_no_conf.csv")
fmp_revenue_2 <- fmp_revenue %>% mutate(Mean = prettyNum(mean_rev, big.mark = ',', scientific = FALSE))
fmp_revenue_3 <- fmp_revenue_2 %>% mutate(Median = prettyNum(med_rev, big.mark = ',', scientific = FALSE))
fmp_revenue_4 <- fmp_revenue_3 %>% mutate(St.Dev = prettyNum(std_rev, big.mark = ',', scientific = FALSE))
fmp_revenue_4$Mean<-paste("$",fmp_revenue_4$Mean)
fmp_revenue_4$Median<-paste("$",fmp_revenue_4$Median)
fmp_revenue_4$St.Dev<-paste("$",fmp_revenue_4$St.Dev)
fmp_revenue_4$Response_Perc<-paste(fmp_revenue_4$Response_Rate,"%")
fmp_revenue_4 <- subset(fmp_revenue_4, select=c(Fishery, Observations, Sampled, Response_Perc, Mean, Median, St.Dev))
knitr::kable(fmp_revenue_4)
fmp_revenue_plot<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/fleets/fmp_boxplots.csv")
fmp_revenue_plot_2 <- ggplot(fmp_revenue_plot, aes(x = Fishery, y = total_revenue, fill=Fishery)) + geom_boxplot(show.legend = TRUE, outlier.shape = NA)
fmp_revenue_plot_2 + coord_cartesian(ylim=c(0, 2000000)) + scale_y_continuous(breaks = seq(0, 2000000, by=250000)) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())

gear_fmp_revenue<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/revenue_summary/gear_fmp_revenue_no_conf.csv")
gear_fmp_revenue_2<- gear_fmp_revenue %>% mutate(Mean = prettyNum(mean_rev, big.mark = ',', scientific = FALSE))
gear_fmp_revenue_3<- gear_fmp_revenue_2 %>% mutate(Median = prettyNum(med_rev, big.mark = ',', scientific = FALSE))
gear_fmp_revenue_4<- gear_fmp_revenue_3 %>% mutate(St.Dev = prettyNum(std_rev, big.mark = ',', scientific = FALSE))
gear_fmp_revenue_4$Mean<-paste("$",gear_fmp_revenue_4$Mean)
gear_fmp_revenue_4$Median<-paste("$",gear_fmp_revenue_4$Median)
gear_fmp_revenue_4$St.Dev<-paste("$",gear_fmp_revenue_4$St.Dev)
gear_fmp_revenue_4$Response_Perc<-paste(gear_fmp_revenue_4$Response_Rate,"%")
gear_fmp_revenue_4<- subset(gear_fmp_revenue_4, select=c(Gear_Fishery, Observations, Sampled, Response_Perc, Mean, Median, St.Dev))
knitr::kable(gear_fmp_revenue_4)
gear_fmp_revenue_plot<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/fleets/gear_fmp_boxplots.csv")
gear_fmp_revenue_plot_2 <- ggplot(gear_fmp_revenue_plot, aes(x = Gear_Fishery, y = total_revenue, fill=Gear_Fishery)) + geom_boxplot(show.legend = TRUE, outlier.shape = NA)
gear_fmp_revenue_plot_2 + coord_cartesian(ylim=c(0, 2000000)) + scale_y_continuous(breaks = seq(0, 2000000, by=250000)) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())
```

By primary gear, scallop dredge and trawl have mean and median values more than twice of those for other gear types. Pot/trap has the third highest median values, followed by gillnet and handgear. For pot/trap and trawl gear, mean values greatly exceed median values, indicating skewness in the distribution to the right. For scallop dredge, the opposite is true.

By FMP, scallop, squid/mackerel/butterfish, and groundfish have the highest mean revenues- all in the $800K-$900K range. However the median value for scallop vessels is considerably higher than all other fisheries. The two highest revenue generating fisheries in the Northeast region during 2022 were lobster and scallop, though the vessel-level mean and medians highlight the differences in these fishing fleets. Median scallop vessel revenues are more than 6x lobster revenues while the number of lobster vessel observations is more than 5x the number of scallop vessels. 

When primary gear and FMP are combined, trawl-groundfish has the highest mean revenue, while scallop dredge-scallop has the highest median revenue. Within primary gear types, a considerable difference in revenues is found across fisheries. For example, mean and median values for lobster pot/trap vessels are roughly twice those values for fluke, black sea bass, and scup pot/trap vessels. Squid trawl vessels have more than twice the mean and median revenues than fluke, black sea bass, and scup trawl vessels.


# Cost Summary
```{r, echo=FALSE}
library(dplyr)
library(scales)

gear_cost<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/cost_summary/gear_cost_no_conf.csv")
gear_cost_2 <- gear_cost %>% mutate(Mean = prettyNum(mean_cost, big.mark = ',', scientific = FALSE))
gear_cost_3 <- gear_cost_2 %>% mutate(Median = prettyNum(med_cost, big.mark = ',', scientific = FALSE))
gear_cost_4 <- gear_cost_3 %>% mutate(St.Dev = prettyNum(std_cost, big.mark = ',', scientific = FALSE))
gear_cost_4$Mean<-paste("$",gear_cost_4$Mean)
gear_cost_4$Median<-paste("$",gear_cost_4$Median)
gear_cost_4$St.Dev<-paste("$",gear_cost_4$St.Dev)
gear_cost_4$Response_Perc<-paste(gear_cost_4$Response_Rate,"%")
gear_cost_4 <- subset(gear_cost_4, select=c(Gear, Observations, Sampled, Response_Perc, Mean, Median, St.Dev))
knitr::kable(gear_cost_4)

fmp_cost<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/cost_summary/fmp_cost_no_conf.csv")
fmp_cost_2 <- fmp_cost %>% mutate(Mean = prettyNum(mean_cost, big.mark = ',', scientific = FALSE))
fmp_cost_3 <- fmp_cost_2 %>% mutate(Median = prettyNum(med_cost, big.mark = ',', scientific = FALSE))
fmp_cost_4 <- fmp_cost_3 %>% mutate(St.Dev = prettyNum(std_cost, big.mark = ',', scientific = FALSE))
fmp_cost_4$Mean<-paste("$",fmp_cost_4$Mean)
fmp_cost_4$Median<-paste("$",fmp_cost_4$Median)
fmp_cost_4$St.Dev<-paste("$",fmp_cost_4$St.Dev)
fmp_cost_4$Response_Perc<-paste(fmp_cost_4$Response_Rate,"%")
fmp_cost_4 <- subset(fmp_cost_4, select=c(Fishery, Observations, Sampled, Response_Perc, Mean, Median, St.Dev))
knitr::kable(fmp_cost_4)

gear_fmp_cost<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/cost_summary/gear_fmp_cost_no_conf.csv")
gear_fmp_cost_2 <- gear_fmp_cost %>% mutate(Mean = prettyNum(mean_cost, big.mark = ',', scientific = FALSE))
gear_fmp_cost_3 <- gear_fmp_cost_2 %>% mutate(Median = prettyNum(med_cost, big.mark = ',', scientific = FALSE))
gear_fmp_cost_4 <- gear_fmp_cost_3 %>% mutate(St.Dev = prettyNum(std_cost, big.mark = ',', scientific = FALSE))
gear_fmp_cost_4$Mean<-paste("$",gear_fmp_cost_4$Mean)
gear_fmp_cost_4$Median<-paste("$",gear_fmp_cost_4$Median)
gear_fmp_cost_4$St.Dev<-paste("$",gear_fmp_cost_4$St.Dev)
gear_fmp_cost_4$Response_Perc<-paste(gear_fmp_cost_4$Response_Rate,"%")
gear_fmp_cost_4 <- subset(gear_fmp_cost_4, select=c(Gear_Fishery, Observations, Sampled, Response_Perc, Mean, Median, St.Dev))
knitr::kable(gear_fmp_cost_4)
```

```{r, echo=FALSE}
library(ggplot2)

#gear_dist<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/cost_contributors/gear_cost_cat_perc.csv")
#knitr::kable(gear_dist)

#ggplot(gear_dist2, aes(x = "", y=Dredge_Scallop, fill = Cost_Category)) + geom_col(color="white") + geom_text(aes(label=percent(Dredge_Scallop)), position = #position_stack(vjust = 0.5), size=6) + coord_polar(theta = "y")  + ggtitle("Dredge_Scallop") + theme(plot.title = element_text(hjust = 0.5, size = 18), 
#axis.title=element_blank(), axis.text=element_blank(), axis.ticks=element_blank(), panel.grid = element_blank(),panel.border = element_blank()) + theme(legend.text = #element_text(size=16), legend.title = element_text(size=16))

gear_dist3<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/cost_contributors/gear_cost_cat_perc_2.csv")
gear_dist3$perc=percent(gear_dist3$percent, accuracy = 0.1)
gear_dist4<- subset(gear_dist3, Gear== "Dredge_Scallop" | Gear== "Gillnet" | Gear== "Handgear")
ggplot(data=gear_dist4, aes(x="", y=percent, fill=Category)) + geom_bar(width = 1, stat = "identity") + coord_polar("y", start=0) + facet_grid(.~Gear) + theme_void() + geom_text(aes(label=perc), position = position_stack(vjust = 0.5), size=3) + theme(legend.text = element_text(size=12), legend.title = element_text(size=12), strip.text.x = element_text(size=11))

gear_dist5<- subset(gear_dist3, Gear== "Longline" | Gear== "Pot/Trap" | Gear== "Seine" | Gear== "Trawl")
ggplot(data=gear_dist5, aes(x="", y=percent, fill=Category)) + geom_bar(width = 1, stat = "identity") + coord_polar("y", start=0) + facet_grid(.~Gear) + theme_void() + geom_text(aes(label=perc), position = position_stack(vjust = 0.5), size=3) + theme(legend.text = element_blank(), legend.title = element_blank(), legend.position = "none", strip.text.x = element_text(size=11))

#gear_dist6<- subset(gear_dist3, Gear== "Trawl")
#ggplot(data=gear_dist6, aes(x="", y=percent, fill=Category)) + geom_bar(width = 1, stat = "identity") + coord_polar("y", start=0) + facet_grid(.~Gear) + theme_void() + #geom_text(aes(label=percent(percent)), position = position_stack(vjust = 0.5), size=4) + theme(legend.text = element_text(size=12), legend.title = element_text(size=12))

#fmp_dist<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/cost_contributors/fmp_cost_cat_perc.csv")
#knitr::kable(fmp_dist)
fmp_dist3<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/cost_contributors/fmp_cost_cat_perc_2.csv")
fmp_dist3$perc=percent(fmp_dist3$percent, accuracy = 0.1)
fmp_dist4<- subset(fmp_dist3, Fishery== "Dogfish" | Fishery== "Fluke, BSB, Scup" | Fishery== "Groundfish")
ggplot(data=fmp_dist4, aes(x="", y=percent, fill=Category)) + geom_bar(width = 1, stat = "identity") + coord_polar("y", start=0) + facet_grid(.~Fishery) + theme_void() + geom_text(aes(label=perc), position = position_stack(vjust = 0.5), size=3) + theme(legend.text = element_text(size=12), legend.title = element_text(size=12), strip.text.x = element_text(size=11))

fmp_dist5<- subset(fmp_dist3, Fishery== "Herring" | Fishery== "Highly Migratory Species" | Fishery== "Lobster" | Fishery== "Monkfish")
ggplot(data=fmp_dist5, aes(x="", y=percent, fill=Category)) + geom_bar(width = 1, stat = "identity") + coord_polar("y", start=0) + facet_grid(.~Fishery) + theme_void() + geom_text(aes(label=perc), position = position_stack(vjust = 0.5), size=3) + theme(legend.text = element_blank(), legend.title = element_blank(), legend.position = "none", strip.text.x = element_text(size=11))

fmp_dist6<- subset(fmp_dist3, Fishery== "Other" | Fishery== "Scallop" | Fishery== "Squid, Mackerel, Butterfish" | Fishery== "Striped Bass")
ggplot(data=fmp_dist6, aes(x="", y=percent, fill=Category)) + geom_bar(width = 1, stat = "identity") + coord_polar("y", start=0) + facet_grid(.~Fishery) + theme_void() + geom_text(aes(label=perc), position = position_stack(vjust = 0.5), size=3) + theme(legend.text = element_blank(), legend.title = element_blank(), legend.position = "none", strip.text.x = element_text(size=11))

#gear_fmp_dist<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/cost_contributors/gear_fmp_cost_cat_perc.csv")
#knitr::kable(gear_fmp_dist)
gear_fmp_dist3<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/cost_contributors/gear_fmp_cost_cat_perc_2.csv")
gear_fmp_dist3$perc=percent(gear_fmp_dist3$percent, accuracy = 0.1)
gear_fmp_dist4<- subset(gear_fmp_dist3, Gear_Fishery== "Dredge_Scallop-Scallop" | Gear_Fishery== "Gillnet-Monkfish" | Gear_Fishery== "Gillnet-Other")
ggplot(data=gear_fmp_dist4, aes(x="", y=percent, fill=Category)) + geom_bar(width = 1, stat = "identity") + coord_polar("y", start=0) + facet_grid(.~Gear_Fishery) + theme_void() + geom_text(aes(label=perc), position = position_stack(vjust = 0.5), size=3) + theme(legend.text = element_text(size=12), legend.title = element_text(size=12), strip.text.x = element_text(size=11))

gear_fmp_dist5<- subset(gear_fmp_dist3, Gear_Fishery== "Handgear-Fluke, BSB, Scup" | Gear_Fishery== "Handgear-Highly Migratory Species" | Gear_Fishery== "Handgear-Other" | Gear_Fishery== "Handgear-Striped Bass")
ggplot(data=gear_fmp_dist5, aes(x="", y=percent, fill=Category)) + geom_bar(width = 1, stat = "identity") + coord_polar("y", start=0) + facet_grid(.~Gear_Fishery) + theme_void() + geom_text(aes(label=perc), position = position_stack(vjust = 0.5), size=3) + theme(legend.text = element_blank(), legend.title = element_blank(), legend.position = "none", strip.text.x = element_text(size=11))

gear_fmp_dist6<- subset(gear_fmp_dist3, Gear_Fishery== "Longline-Highly Migratory Species" | Gear_Fishery== "Pot/Trap-Fluke, BSB, Scup" | Gear_Fishery== "Pot/Trap-Lobster" | Gear_Fishery== "Pot/Trap-Other")
ggplot(data=gear_fmp_dist6, aes(x="", y=percent, fill=Category)) + geom_bar(width = 1, stat = "identity") + coord_polar("y", start=0) + facet_grid(.~Gear_Fishery) + theme_void() + geom_text(aes(label=perc), position = position_stack(vjust = 0.5), size=3) + theme(legend.text = element_blank(), legend.title = element_blank(), legend.position = "none", strip.text.x = element_text(size=11))

gear_fmp_dist7<- subset(gear_fmp_dist3, Gear_Fishery== "Trawl-Fluke, BSB, Scup" | Gear_Fishery== "Trawl-Groundfish" | Gear_Fishery== "Trawl-Other" | Gear_Fishery== "Trawl-Squid, Mackerel, Butterfish")
ggplot(data=gear_fmp_dist7, aes(x="", y=percent, fill=Category)) + geom_bar(width = 1, stat = "identity") + coord_polar("y", start=0) + facet_grid(.~Gear_Fishery) + theme_void() + geom_text(aes(label=perc), position = position_stack(vjust = 0.5), size=3) + theme(legend.text = element_blank(), legend.title = element_blank(), legend.position = "none", strip.text.x = element_text(size=11))
```

By primary gear, the order of costs from highest to lowest follows the same order as with revenue- scallop dredge, trawl, pot/trap, gillnet, handgear. Scallop dredge and trawl again have mean and median values more than twice of those for other gear types. These highest cost gear types are crew-intensive with crew expenses accounting for over 40% of total costs among scallop dredge and trawl vessels. Pot/trap and gillnet vessels have the highest share of trip expenses. For pot/trap vessels, high trip expenses can be explained in part by the cost of bait. For handgear vessels, the lowest cost gear group, repair/maintenance/upgrade/improvement (RMUI) is the largest share of costs. Overhead costs as a share of total are similar across primary gear groups, ranging from 6-10%. Quota net costs are not included in the figures as they represent a relatively small portion of total costs; the highest share was 4% for the longline group. Some gear groups have negative quota net costs, in which leasing revenue exceeds costs. However, for some individual vessels, quota leasing may represent a substantial cost.

By FMP, a number of fisheries have similar mean costs. Scallop vessels have the highest mean costs, though groundfish herring, and squid, mackerel, butterfish all have mean values within 5% of scallop vessels. The squid, mackerel, butterfish and scallop fisheries are the most crew-intensive, with crew payments exceeding 40% of total costs for both fisheries. The lobster fishery has relatively higher trip expenses than most other FMPs, with 28% of total costs in this category. Quota net costs are not included in the figures as they represent a relatively small portion of total costs; the highest share was 4% for groundfish. Some FMP groups have negative quota net costs, in which leasing revenue exceeds costs. However, for some individual vessels, quota leasing may represent a substantial cost.

When primary gear and FMP are combined, trawl-groundfish has the highest mean and median costs. Within primary gear types, a considerable difference in costs is found across fisheries. For example, mean and median values for lobster pot/trap vessels are greater than those values for fluke, black sea bass, and scup pot/trap vessels. The top cost contributors are generally similar within a gear group and across fisheries. For example, all trawl fisheries have relatively high crew payments, ~40% of total costs. Handgear fisheries have relatively large shares of repair/maintenance/upgrade/improvement (RMUI) costs. Pot/trap fisheries have relatively high trip expenses ~25-30% of total costs, possibly due to the cost of bait. Quota net costs are not included in the figures as they represent a relatively small portion of total costs; the highest share was 4% for Longline-HMS and Trawl- groundfish. Some gear/FMP groups have negative quota net costs, in which leasing revenue exceeds costs. However, for some individual vessels, quota leasing may represent a substantial cost.


# Operating Profit
```{r, echo=FALSE}
library(ggplot2)

gear_op_stats<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/profit_summary/summary_stats_op_gear.csv")
gear_op_stats_2 <- gear_op_stats %>% mutate(Mean = prettyNum(mean_op, big.mark = ',', scientific = FALSE))
gear_op_stats_3 <- gear_op_stats_2 %>% mutate(Median = prettyNum(med_op, big.mark = ',', scientific = FALSE))
gear_op_stats_4 <- gear_op_stats_3 %>% mutate(St.Dev = prettyNum(std_op, big.mark = ',', scientific = FALSE))
gear_op_stats_4$Mean<-paste("$",gear_op_stats_4$Mean)
gear_op_stats_4$Median<-paste("$",gear_op_stats_4$Median)
gear_op_stats_4$St.Dev<-paste("$",gear_op_stats_4$St.Dev)
gear_op_stats_4$Response_Perc<-paste(gear_op_stats_4$Response_Rate,"%")
gear_op_stats_4 <- subset(gear_op_stats_4, select=c(Gear, Observations, Sampled, Response_Perc, Mean, Median, St.Dev))
knitr::kable(gear_op_stats_4)
gear_profit<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/fleets/gear_boxplots.csv")
gear_op <- ggplot(gear_profit, aes(x = Gear, y = total_operating_profit, fill=Gear)) + geom_boxplot(show.legend = FALSE, outlier.shape = NA)
gear_op + coord_cartesian(ylim=c(-250000, 1500000)) + scale_y_continuous(breaks = seq(-250000, 1500000, by=250000))

fmp_op_stats<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/profit_summary/summary_stats_op_fmp.csv")
fmp_op_stats_2 <- fmp_op_stats %>% mutate(Mean = prettyNum(mean_op, big.mark = ',', scientific = FALSE))
fmp_op_stats_3 <- fmp_op_stats_2 %>% mutate(Median = prettyNum(med_op, big.mark = ',', scientific = FALSE))
fmp_op_stats_4 <- fmp_op_stats_3 %>% mutate(St.Dev = prettyNum(std_op, big.mark = ',', scientific = FALSE))
fmp_op_stats_4$Mean<-paste("$",fmp_op_stats_4$Mean)
fmp_op_stats_4$Median<-paste("$",fmp_op_stats_4$Median)
fmp_op_stats_4$St.Dev<-paste("$",fmp_op_stats_4$St.Dev)
fmp_op_stats_4$Response_Perc<-paste(fmp_op_stats_4$Response_Rate,"%")
fmp_op_stats_4 <- subset(fmp_op_stats_4, select=c(Fishery, Observations, Sampled, Response_Perc, Mean, Median, St.Dev))
knitr::kable(fmp_op_stats_4)
fmp_profit<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/fleets/fmp_boxplots.csv")
fmp_op <- ggplot(fmp_profit, aes(x = Fishery, y = total_operating_profit, fill=Fishery)) + geom_boxplot(show.legend = TRUE, outlier.shape = NA)
fmp_op + coord_cartesian(ylim=c(-250000, 1500000)) + scale_y_continuous(breaks = seq(-250000, 1500000, by=250000)) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())

gear_fmp_op_stats<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/profit_summary/summary_stats_op_gear_fmp.csv")
gear_fmp_op_stats_2 <- gear_fmp_op_stats %>% mutate(Mean = prettyNum(mean_op, big.mark = ',', scientific = FALSE))
gear_fmp_op_stats_3 <- gear_fmp_op_stats_2 %>% mutate(Median = prettyNum(med_op, big.mark = ',', scientific = FALSE))
gear_fmp_op_stats_4 <- gear_fmp_op_stats_3 %>% mutate(St.Dev = prettyNum(std_op, big.mark = ',', scientific = FALSE))
gear_fmp_op_stats_4$Mean<-paste("$",gear_fmp_op_stats_4$Mean)
gear_fmp_op_stats_4$Median<-paste("$",gear_fmp_op_stats_4$Median)
gear_fmp_op_stats_4$St.Dev<-paste("$",gear_fmp_op_stats_4$St.Dev)
gear_fmp_op_stats_4$Response_Perc<-paste(gear_fmp_op_stats_4$Response_Rate,"%")
gear_fmp_op_stats_4 <- subset(gear_fmp_op_stats_4, select=c(Gear_Fishery, Observations, Sampled, Response_Perc, Mean, Median, St.Dev))
knitr::kable(gear_fmp_op_stats_4)
gear_fmp_profit<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/fleets/gear_fmp_boxplots.csv")
gear_fmp_op <- ggplot(gear_fmp_profit, aes(x = Gear_Fishery, y = total_operating_profit, fill=Gear_Fishery)) + geom_boxplot(show.legend = TRUE, outlier.shape = NA)
gear_fmp_op + coord_cartesian(ylim=c(-250000, 1500000)) + scale_y_continuous(breaks = seq(-250000, 1500000, by=250000)) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())
```

By primary gear, scallop dredge has considerably higher mean and median operating profit than other gear groups. Trawl has the second highest mean and median values, followed by pot/trap, gillnet, and handgear. For most gear types, mean values greatly exceed median values, indicating skewness in the distribution to the right. The boxplots emphasize the range of operating profit by gear group with the box extending past $750K for scallop dredge and ~$500k for trawl. Pot/trap and gillnet extend considerably less and handgear is close to $0. A small number of observations exhibit negative operating profit, which generally is viewed as the short-term shutdown point in which continuing operations is no longer economically viable. These observations are likely explained in part by other vessel uses outside of commercial fishing. For example, if a vessel is used for leisure activities, the vessel owner may have listed the costs of vessel maintenance and upkeep for activities which do not generate revenue. These vessels tend to belong to the Handgear and Pot/Trap gear groups.

By FMP, scallop has higher mean and median operating profit than all other fisheries. Squid, mackerel, butterfish has the second highest mean values, while herring has the second highest median values. Groundfish has the third highest mean values. For a number of fisheries, mean values greatly exceed median values, indicating skewness in the distribution to the right. For lobster and scallop, the difference between mean and median values is relatively small. The boxplots emphasize the range of operating profit by gear group with the box extending past $750K for scallop dredge and close to $500k for groundfish, herring, and squid, mackerel, butterfish. Lobster extends to ~100k and other fisheries are close to $0.  A small number of observations exhibit negative operating profit, which generally is viewed as the shutdown point in which continuing operations is no longer economically viable. These observations are likely explained in part by other vessel uses outside of commercial fishing. For example, if a vessel is used for leisure activities, the vessel owner may have listed the costs of vessel maintenance and upkeep for activities which do not generate revenue. Many of these vessels belong to the Highly Migratory Species fishery.

When primary gear and FMP are combined, trawl-groundfish has the highest mean operating profit, while scallop dredge-scallop has the highest median value. Within primary gear types, a considerable difference in operating profit is found across fisheries. For example, the mean value for squid trawl vessels is roughly twice that for fluke, black sea bass, and scup trawl vessels. Within the pot/trap gear group, summer flounder, black sea bass, scup vessels have a higher median value than for lobster but a lower median value. A small number of observations exhibit negative operating profit, which generally is viewed as the shutdown point in which continuing operations is no longer economically viable. These observations are likely explained in part by other vessel uses outside of commercial fishing. For example, if a vessel is used for leisure activities, the vessel owner may have listed the costs of vessel maintenance and upkeep for activities which do not generate revenue.


# Total Profit
```{r, echo=FALSE}
library(ggplot2)

gear_tp_stats<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/profit_summary/summary_stats_tp_gear.csv")
gear_tp_stats_2 <- gear_tp_stats %>% mutate(Mean = prettyNum(mean_tp, big.mark = ',', scientific = FALSE))
gear_tp_stats_3 <- gear_tp_stats_2 %>% mutate(Median = prettyNum(med_tp, big.mark = ',', scientific = FALSE))
gear_tp_stats_4 <- gear_tp_stats_3 %>% mutate(St.Dev = prettyNum(std_tp, big.mark = ',', scientific = FALSE))
gear_tp_stats_4$Mean<-paste("$",gear_tp_stats_4$Mean)
gear_tp_stats_4$Median<-paste("$",gear_tp_stats_4$Median)
gear_tp_stats_4$St.Dev<-paste("$",gear_tp_stats_4$St.Dev)
gear_tp_stats_4$Response_Perc<-paste(gear_tp_stats_4$Response_Rate,"%")
gear_tp_stats_4 <- subset(gear_tp_stats_4, select=c(Gear, Observations, Sampled, Response_Perc, Mean, Median, St.Dev))
knitr::kable(gear_tp_stats_4)
gear_tp <- ggplot(gear_profit, aes(x = Gear, y = total_profit, fill=Gear)) + geom_boxplot(show.legend = FALSE, outlier.shape = NA)
gear_tp + coord_cartesian(ylim=c(-500000, 1000000)) + scale_y_continuous(breaks = seq(-500000, 1000000, by=250000))

fmp_tp_stats<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/profit_summary/summary_stats_tp_fmp.csv")
fmp_tp_stats_2 <- fmp_tp_stats %>% mutate(Mean = prettyNum(mean_tp, big.mark = ',', scientific = FALSE))
fmp_tp_stats_3 <- fmp_tp_stats_2 %>% mutate(Median = prettyNum(med_tp, big.mark = ',', scientific = FALSE))
fmp_tp_stats_4 <- fmp_tp_stats_3 %>% mutate(St.Dev = prettyNum(std_tp, big.mark = ',', scientific = FALSE))
fmp_tp_stats_4$Mean<-paste("$",fmp_tp_stats_4$Mean)
fmp_tp_stats_4$Median<-paste("$",fmp_tp_stats_4$Median)
fmp_tp_stats_4$St.Dev<-paste("$",fmp_tp_stats_4$St.Dev)
fmp_tp_stats_4$Response_Perc<-paste(fmp_tp_stats_4$Response_Rate,"%")
fmp_tp_stats_4 <- subset(fmp_tp_stats_4, select=c(Fishery, Observations, Sampled, Response_Perc, Mean, Median, St.Dev))
knitr::kable(fmp_tp_stats_4)
fmp_tp <- ggplot(fmp_profit, aes(x = Fishery, y = total_profit, fill=Fishery)) + geom_boxplot(show.legend = TRUE, outlier.shape = NA)
fmp_tp + coord_cartesian(ylim=c(-500000, 1000000)) + scale_y_continuous(breaks = seq(-500000, 1000000, by=250000)) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())

gear_fmp_tp_stats<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/profit_summary/summary_stats_tp_gear_fmp.csv")
gear_fmp_tp_stats_2 <- gear_fmp_tp_stats %>% mutate(Mean = prettyNum(mean_tp, big.mark = ',', scientific = FALSE))
gear_fmp_tp_stats_3 <- gear_fmp_tp_stats_2 %>% mutate(Median = prettyNum(med_tp, big.mark = ',', scientific = FALSE))
gear_fmp_tp_stats_4 <- gear_fmp_tp_stats_3 %>% mutate(St.Dev = prettyNum(std_tp, big.mark = ',', scientific = FALSE))
gear_fmp_tp_stats_4$Mean<-paste("$",gear_fmp_tp_stats_4$Mean)
gear_fmp_tp_stats_4$Median<-paste("$",gear_fmp_tp_stats_4$Median)
gear_fmp_tp_stats_4$St.Dev<-paste("$",gear_fmp_tp_stats_4$St.Dev)
gear_fmp_tp_stats_4$Response_Perc<-paste(gear_fmp_tp_stats_4$Response_Rate,"%")
gear_fmp_tp_stats_4 <- subset(gear_fmp_tp_stats_4, select=c(Gear_Fishery, Observations, Sampled, Response_Perc, Mean, Median, St.Dev))
knitr::kable(gear_fmp_tp_stats_4)
gear_fmp_tp <- ggplot(gear_fmp_profit, aes(x = Gear_Fishery, y = total_profit, fill=Gear_Fishery)) + geom_boxplot(show.legend = TRUE, outlier.shape = NA)
gear_fmp_tp + coord_cartesian(ylim=c(-500000, 1000000)) + scale_y_continuous(breaks = seq(-500000, 1000000, by=250000)) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())
```

By primary gear, scallop dredge has considerably higher mean and median total profit than other gear groups. Trawl has the second highest mean and median values. Gillnet, handgear, and pot/trap show negative mean profit while the median value is slightly positive for pot/trap. For scallop dredge and trawl gear, mean values greatly exceed median values. For pot/trap the opposite is true. A considerable range of profits is evident in all gear groups, especially scallop dredge and trawl. These wide ranges are to be expected as some vessels have extensive repairs and/or upgrades during 2022 while other vessels may have incurred these costs in other years. If a vessel were to show negative total profit in the long run, vessel operations would not be economically viable. Since the cost survey is not conducted on an annual basis, the ability to assess long-term vessel profits is limited.

By FMP, scallop and squid, mackerel butterfish have the highest mean and median total profits. By median values, the difference between these two fisheries and all other FMPs is considerable. Groundfish has the third highest mean and median values. The lobster fleet, representing over half of the total observations, shows slightly positive median profits. The fluke black sea bass, scup fishery shows similar limited median operating profit as the lobster fishery. A considerable range of profits is evident in many fisheries. These wide ranges are to be expected as some vessels have extensive repairs and/or upgrades during 2022 while other vessels may have incurred these costs in other years. If a vessel were to show negative total profit in the long run, vessel operations would not be economically viable. Since the cost survey is not conducted on an annual basis, the ability to assess long-term vessel profits is limited.

When primary gear and FMP are combined, trawl-squid, mackerel, butterfish, trawl-groundfish, and scallop dredge-scallop have the highest mean profits. Median values tell a different story, with trawl groundfish showing profits more similar to trawl vessels in the summer flounder, black sea bass, scup fishery. Median profits are highest for trawl-squid, mackerel, butterfish, and scallop dredge-scallop. Within primary gear types, a considerable difference in total profit is found across fisheries. For example, within the pot/trap gear group, summer flounder, black sea bass, scup vessels have higher median profits compared to lobster. This is noteworthy as lobster vessels showed higher revenues than summer flounder, black sea bass, scup vessels within the pot/trap gear group. Within the Trawl group, squid, mackerel, butterfish vessels show substantially higher median profits compared to groundfish or summer flounder, black sea bass, scup vessels. Wide ranges within Gear/FMP groupings are to be expected as some vessels have extensive repairs and/or upgrades during 2022 while other vessels may have incurred these costs in other years. If a vessel were to show negative total profit in the long run, vessel operations would not be economically viable. Since the cost survey is not conducted on an annual basis, the ability to assess long-term vessel profits is limited.


# Discussion

## Applications
These profit profiles are intended to support the management of commercial fisheries in the Northeast U.S. The results presented have a number of different applications, including:

1. Allowing end-users to gain a better understanding of cost distributions by fishing fleet. In economic analyses to support management, operating profit/net revenue is      most often calculated. Across FMPs, operating costs account for varying levels of total costs.
2. Allowing end-users to gain a better understanding of how gross revenues translate to profit by fishing fleet.  
3. Supporting research into the viability of fishing operations. By examining the economic health of commercial vessels, the implications of management actions on fisheries  can be better understood. These potential actions include, but are not limited to:
    a. Changes in fishing quotas/days-at-sea 
    b. Re-allocation of quota among sectors (e.g. commercial vs. recreational)
    c. Establishment or modification of accumulation limits
    d. Changes in costs faced by fishing vessels (e.g. monitoring) 

For management purposes, these profiles can be incorporated into the Affected Environment: Human Communities section for various FMP actions. This section has frequently focused on trends in gross revenue and fishery effort. Profit provides additional information in capturing the current state of the fishery. When looking at operating profit, mean and median values are relatively close for a number of gear types (scallop dredge, longline, pot trap) and fisheries (herring, lobster, scallop). The disparity between mean and median tends to increase when looking at total profit as more sporadic costs are incorporated. In other words, it is difficult to assess what a “typical vessel” in a particular gear group or fishery looks given large distributions. 

The profiles can also be incorporated into the economic impact analysis of management actions. For example, the ratio of total profits to gross revenues can inform how changes in quota impact the bottom line of vessel operators. For the squid, mackerel, butterfish fishery, the ratio of total profit/gross revenue is 33% when using mean values and 26% using median values. For the scallop fishery, these numbers are 32% and 17% respectively. For the groundfish fishery, these numbers are 26% and 8%.

## Limitations and Future Work
The use of these profit profiles does have some limitations. The Commercial Fishing Business Cost Survey is conducted within the Northeast region every few years. For some vessel owners, 2022 may have been an unusually profitable year, either through increased revenues, decreased costs, or a combination. In contrast, other vessels likely experienced a less profitable year than normal in 2022. A vessel owner may have been more/less likely to respond based on their economic situation during 2022. Since the cost survey is not conducted on an annual basis, and the individuals that respond to the survey vary, it is difficult to assess this effect.  

As more years of data are available, the better we can understand the economic performance of fishing fleets in the region. Data collected in future surveys will feed into the profit calculations presented on this site. Similar to the gear/FMP dropdown function, the user will be able to query the year of interest.


# References

Ardini G., Murphy T., Werner S., Bailey M. 2022. An Overview of the Social Sciences Branch (SSB) Commercial Fishing Business Cost Survey in the Northeast: Protocol and Results for Survey Years 2011, 2012, and 2015. US Dept Commer, Northeast Fish Sci Cent Tech Memo 278. 146 p.

Das C. 2013. An Overview of the Annual Cost Survey Protocol and Results in the Northeast (2007 to 2009). US Dept Commer, Northeast Fish Sci Cent Tech Memo 226. 45 p.

Kitts A., Walden J., Squires D., Travis M., Steiner E., Pfeiffer L., Liese C., Pan M. 2022. NOAA Fisheries Best Practices for Measuring Returns to Fishing Businesses. US Dept Commer, NOAA Tech Memo NMFS-F/SPO-231, 54 p.

(forthcoming) Werner S., Conley E., Ardini G. 2025. Greater Atlantic Region Commercial Fishing Business Cost Survey for 2022: Survey Implementation, Data Processing Methods and Preliminary Results.

