# Operating Profit
```{r, echo=FALSE}
library(ggplot2)

gear_op_stats<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/profit_summary/summary_stats_op_gear.csv")
gear_op_stats_2 <- gear_op_stats %>% mutate(Mean = prettyNum(mean_op, big.mark = ',', scientific = FALSE))
gear_op_stats_3 <- gear_op_stats_2 %>% mutate(Median = prettyNum(med_op, big.mark = ',', scientific = FALSE))
gear_op_stats_4 <- gear_op_stats_3 %>% mutate(St.Dev = prettyNum(std_op, big.mark = ',', scientific = FALSE))
gear_op_stats_4$Mean<-paste("$",gear_op_stats_4$Mean)
gear_op_stats_4$Median<-paste("$",gear_op_stats_4$Median)
gear_op_stats_4$St.Dev<-paste("$",gear_op_stats_4$St.Dev)
gear_op_stats_4$Response_Perc<-paste(gear_op_stats_4$Response_Rate,"%")
gear_op_stats_4 <- subset(gear_op_stats_4, select=c(Gear, Observations, Sampled, Response_Perc, Mean, Median, St.Dev))
knitr::kable(gear_op_stats_4)
gear_profit<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/fleets/gear_boxplots.csv")
gear_op <- ggplot(gear_profit, aes(x = Gear, y = total_operating_profit, fill=Gear)) + geom_boxplot(show.legend = FALSE, outlier.shape = NA)
gear_op + coord_cartesian(ylim=c(-250000, 1500000)) + scale_y_continuous(breaks = seq(-250000, 1500000, by=250000))

fmp_op_stats<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/profit_summary/summary_stats_op_fmp.csv")
fmp_op_stats_2 <- fmp_op_stats %>% mutate(Mean = prettyNum(mean_op, big.mark = ',', scientific = FALSE))
fmp_op_stats_3 <- fmp_op_stats_2 %>% mutate(Median = prettyNum(med_op, big.mark = ',', scientific = FALSE))
fmp_op_stats_4 <- fmp_op_stats_3 %>% mutate(St.Dev = prettyNum(std_op, big.mark = ',', scientific = FALSE))
fmp_op_stats_4$Mean<-paste("$",fmp_op_stats_4$Mean)
fmp_op_stats_4$Median<-paste("$",fmp_op_stats_4$Median)
fmp_op_stats_4$St.Dev<-paste("$",fmp_op_stats_4$St.Dev)
fmp_op_stats_4$Response_Perc<-paste(fmp_op_stats_4$Response_Rate,"%")
fmp_op_stats_4 <- subset(fmp_op_stats_4, select=c(Fishery, Observations, Sampled, Response_Perc, Mean, Median, St.Dev))
knitr::kable(fmp_op_stats_4)
fmp_profit<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/fleets/fmp_boxplots.csv")
fmp_op <- ggplot(fmp_profit, aes(x = Fishery, y = total_operating_profit, fill=Fishery)) + geom_boxplot(show.legend = TRUE, outlier.shape = NA)
fmp_op + coord_cartesian(ylim=c(-250000, 1500000)) + scale_y_continuous(breaks = seq(-250000, 1500000, by=250000)) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())

gear_fmp_op_stats<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/profit_summary/summary_stats_op_gear_fmp.csv")
gear_fmp_op_stats_2 <- gear_fmp_op_stats %>% mutate(Mean = prettyNum(mean_op, big.mark = ',', scientific = FALSE))
gear_fmp_op_stats_3 <- gear_fmp_op_stats_2 %>% mutate(Median = prettyNum(med_op, big.mark = ',', scientific = FALSE))
gear_fmp_op_stats_4 <- gear_fmp_op_stats_3 %>% mutate(St.Dev = prettyNum(std_op, big.mark = ',', scientific = FALSE))
gear_fmp_op_stats_4$Mean<-paste("$",gear_fmp_op_stats_4$Mean)
gear_fmp_op_stats_4$Median<-paste("$",gear_fmp_op_stats_4$Median)
gear_fmp_op_stats_4$St.Dev<-paste("$",gear_fmp_op_stats_4$St.Dev)
gear_fmp_op_stats_4$Response_Perc<-paste(gear_fmp_op_stats_4$Response_Rate,"%")
gear_fmp_op_stats_4 <- subset(gear_fmp_op_stats_4, select=c(Gear_Fishery, Observations, Sampled, Response_Perc, Mean, Median, St.Dev))
knitr::kable(gear_fmp_op_stats_4)
gear_fmp_profit<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/fleets/gear_fmp_boxplots.csv")
gear_fmp_op <- ggplot(gear_fmp_profit, aes(x = Gear_Fishery, y = total_operating_profit, fill=Gear_Fishery)) + geom_boxplot(show.legend = TRUE, outlier.shape = NA)
gear_fmp_op + coord_cartesian(ylim=c(-250000, 1500000)) + scale_y_continuous(breaks = seq(-250000, 1500000, by=250000)) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())
```

By primary gear, scallop dredge has considerably higher mean and median operating profit than other gear groups. Trawl has the second highest mean and median values, followed by pot/trap, gillnet, and handgear. For most gear types, mean values greatly exceed median values, indicating skewness in the distribution to the right. The boxplots emphasize the range of operating profit by gear group with the box extending past $750K for scallop dredge and ~$500k for trawl. Pot/trap and gillnet extend considerably less and handgear is close to $0. A small number of observations exhibit negative operating profit, which generally is viewed as the short-term shutdown point in which continuing operations is no longer economically viable. These observations are likely explained in part by other vessel uses outside of commercial fishing. For example, if a vessel is used for leisure activities, the vessel owner may have listed the costs of vessel maintenance and upkeep for activities which do not generate revenue. These vessels tend to belong to the Handgear and Pot/Trap gear groups.

By FMP, scallop has higher mean and median operating profit than all other fisheries. Squid, mackerel, butterfish has the second highest mean values, while herring has the second highest median values. Groundfish has the third highest mean values. For a number of fisheries, mean values greatly exceed median values, indicating skewness in the distribution to the right. For lobster and scallop, the difference between mean and median values is relatively small. The boxplots emphasize the range of operating profit by gear group with the box extending past $750K for scallop dredge and close to $500k for groundfish, herring, and squid, mackerel, butterfish. Lobster extends to ~100k and other fisheries are close to $0.  A small number of observations exhibit negative operating profit, which generally is viewed as the shutdown point in which continuing operations is no longer economically viable. These observations are likely explained in part by other vessel uses outside of commercial fishing. For example, if a vessel is used for leisure activities, the vessel owner may have listed the costs of vessel maintenance and upkeep for activities which do not generate revenue. Many of these vessels belong to the Highly Migratory Species fishery.

When primary gear and FMP are combined, trawl-groundfish has the highest mean operating profit, while scallop dredge-scallop has the highest median value. Within primary gear types, a considerable difference in operating profit is found across fisheries. For example, the mean value for squid trawl vessels is roughly twice that for fluke, black sea bass, and scup trawl vessels. Within the pot/trap gear group, summer flounder, black sea bass, scup vessels have a higher median value than for lobster but a lower median value. A small number of observations exhibit negative operating profit, which generally is viewed as the shutdown point in which continuing operations is no longer economically viable. These observations are likely explained in part by other vessel uses outside of commercial fishing. For example, if a vessel is used for leisure activities, the vessel owner may have listed the costs of vessel maintenance and upkeep for activities which do not generate revenue.


# Total Profit
```{r, echo=FALSE}
library(ggplot2)

gear_tp_stats<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/profit_summary/summary_stats_tp_gear.csv")
gear_tp_stats_2 <- gear_tp_stats %>% mutate(Mean = prettyNum(mean_tp, big.mark = ',', scientific = FALSE))
gear_tp_stats_3 <- gear_tp_stats_2 %>% mutate(Median = prettyNum(med_tp, big.mark = ',', scientific = FALSE))
gear_tp_stats_4 <- gear_tp_stats_3 %>% mutate(St.Dev = prettyNum(std_tp, big.mark = ',', scientific = FALSE))
gear_tp_stats_4$Mean<-paste("$",gear_tp_stats_4$Mean)
gear_tp_stats_4$Median<-paste("$",gear_tp_stats_4$Median)
gear_tp_stats_4$St.Dev<-paste("$",gear_tp_stats_4$St.Dev)
gear_tp_stats_4$Response_Perc<-paste(gear_tp_stats_4$Response_Rate,"%")
gear_tp_stats_4 <- subset(gear_tp_stats_4, select=c(Gear, Observations, Sampled, Response_Perc, Mean, Median, St.Dev))
knitr::kable(gear_tp_stats_4)
gear_tp <- ggplot(gear_profit, aes(x = Gear, y = total_profit, fill=Gear)) + geom_boxplot(show.legend = FALSE, outlier.shape = NA)
gear_tp + coord_cartesian(ylim=c(-500000, 1000000)) + scale_y_continuous(breaks = seq(-500000, 1000000, by=250000))

fmp_tp_stats<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/profit_summary/summary_stats_tp_fmp.csv")
fmp_tp_stats_2 <- fmp_tp_stats %>% mutate(Mean = prettyNum(mean_tp, big.mark = ',', scientific = FALSE))
fmp_tp_stats_3 <- fmp_tp_stats_2 %>% mutate(Median = prettyNum(med_tp, big.mark = ',', scientific = FALSE))
fmp_tp_stats_4 <- fmp_tp_stats_3 %>% mutate(St.Dev = prettyNum(std_tp, big.mark = ',', scientific = FALSE))
fmp_tp_stats_4$Mean<-paste("$",fmp_tp_stats_4$Mean)
fmp_tp_stats_4$Median<-paste("$",fmp_tp_stats_4$Median)
fmp_tp_stats_4$St.Dev<-paste("$",fmp_tp_stats_4$St.Dev)
fmp_tp_stats_4$Response_Perc<-paste(fmp_tp_stats_4$Response_Rate,"%")
fmp_tp_stats_4 <- subset(fmp_tp_stats_4, select=c(Fishery, Observations, Sampled, Response_Perc, Mean, Median, St.Dev))
knitr::kable(fmp_tp_stats_4)
fmp_tp <- ggplot(fmp_profit, aes(x = Fishery, y = total_profit, fill=Fishery)) + geom_boxplot(show.legend = TRUE, outlier.shape = NA)
fmp_tp + coord_cartesian(ylim=c(-500000, 1000000)) + scale_y_continuous(breaks = seq(-500000, 1000000, by=250000)) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())

gear_fmp_tp_stats<- read.csv(file="V:/Ardini_Cost_Survey/SAS_Cost_Survey/Profitability_Profiles/Calculate_Profit/Final/profit_summary/summary_stats_tp_gear_fmp.csv")
gear_fmp_tp_stats_2 <- gear_fmp_tp_stats %>% mutate(Mean = prettyNum(mean_tp, big.mark = ',', scientific = FALSE))
gear_fmp_tp_stats_3 <- gear_fmp_tp_stats_2 %>% mutate(Median = prettyNum(med_tp, big.mark = ',', scientific = FALSE))
gear_fmp_tp_stats_4 <- gear_fmp_tp_stats_3 %>% mutate(St.Dev = prettyNum(std_tp, big.mark = ',', scientific = FALSE))
gear_fmp_tp_stats_4$Mean<-paste("$",gear_fmp_tp_stats_4$Mean)
gear_fmp_tp_stats_4$Median<-paste("$",gear_fmp_tp_stats_4$Median)
gear_fmp_tp_stats_4$St.Dev<-paste("$",gear_fmp_tp_stats_4$St.Dev)
gear_fmp_tp_stats_4$Response_Perc<-paste(gear_fmp_tp_stats_4$Response_Rate,"%")
gear_fmp_tp_stats_4 <- subset(gear_fmp_tp_stats_4, select=c(Gear_Fishery, Observations, Sampled, Response_Perc, Mean, Median, St.Dev))
knitr::kable(gear_fmp_tp_stats_4)
gear_fmp_tp <- ggplot(gear_fmp_profit, aes(x = Gear_Fishery, y = total_profit, fill=Gear_Fishery)) + geom_boxplot(show.legend = TRUE, outlier.shape = NA)
gear_fmp_tp + coord_cartesian(ylim=c(-500000, 1000000)) + scale_y_continuous(breaks = seq(-500000, 1000000, by=250000)) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())
```

By primary gear, scallop dredge has considerably higher mean and median total profit than other gear groups. Trawl has the second highest mean and median values. Gillnet, handgear, and pot/trap show negative mean profit while the median value is slightly positive for pot/trap. For scallop dredge and trawl gear, mean values greatly exceed median values. For pot/trap the opposite is true. A considerable range of profits is evident in all gear groups, especially scallop dredge and trawl. These wide ranges are to be expected as some vessels have extensive repairs and/or upgrades during 2022 while other vessels may have incurred these costs in other years. If a vessel were to show negative total profit in the long run, vessel operations would not be economically viable. Since the cost survey is not conducted on an annual basis, the ability to assess long-term vessel profits is limited.

By FMP, scallop and squid, mackerel butterfish have the highest mean and median total profits. By median values, the difference between these two fisheries and all other FMPs is considerable. Groundfish has the third highest mean and median values. The lobster fleet, representing over half of the total observations, shows slightly positive median profits. The fluke black sea bass, scup fishery shows similar limited median operating profit as the lobster fishery. A considerable range of profits is evident in many fisheries. These wide ranges are to be expected as some vessels have extensive repairs and/or upgrades during 2022 while other vessels may have incurred these costs in other years. If a vessel were to show negative total profit in the long run, vessel operations would not be economically viable. Since the cost survey is not conducted on an annual basis, the ability to assess long-term vessel profits is limited.

When primary gear and FMP are combined, trawl-squid, mackerel, butterfish, trawl-groundfish, and scallop dredge-scallop have the highest mean profits. Median values tell a different story, with trawl groundfish showing profits more similar to trawl vessels in the summer flounder, black sea bass, scup fishery. Median profits are highest for trawl-squid, mackerel, butterfish, and scallop dredge-scallop. Within primary gear types, a considerable difference in total profit is found across fisheries. For example, within the pot/trap gear group, summer flounder, black sea bass, scup vessels have higher median profits compared to lobster. This is noteworthy as lobster vessels showed higher revenues than summer flounder, black sea bass, scup vessels within the pot/trap gear group. Within the Trawl group, squid, mackerel, butterfish vessels show substantially higher median profits compared to groundfish or summer flounder, black sea bass, scup vessels. Wide ranges within Gear/FMP groupings are to be expected as some vessels have extensive repairs and/or upgrades during 2022 while other vessels may have incurred these costs in other years. If a vessel were to show negative total profit in the long run, vessel operations would not be economically viable. Since the cost survey is not conducted on an annual basis, the ability to assess long-term vessel profits is limited.
